CUDA-accelerated genetic feedforward-ANN training for data mining

### 2010-07-10 doc/HPCS-proceedings-paper.tex
Rephrase in conclusion why multipass was required
git-svn-id: http://svn.scesoc.ca/evolution/trunk@129 fee98b4c-d1c3-4d34-9525-a468159e4a48
Reword how the top 20% counting is done
git-svn-id: http://svn.scesoc.ca/evolution/trunk@130 fee98b4c-d1c3-4d34-9525-a468159e4a48

% ----------------------------------------------------------------------------
\subsection{Computation of Number of Truly Positive Top 20\% Instances} \label{truepos}
% ----------------------------------------------------------------------------
To compute the number of truly positive instances in the top 20\% of scores, if training instances are unsorted, we must read from memory each instance's true class and corresponding ANN output value and compare the output value to the candidate's threshold. For $n$ training instances, this requires $2 * n$ memory accesses. However, by keeping the $n_p$ positive instances first in memory, we need only read the first $n_p$ ANN outputs, knowing that they correspond to the positive instances. This reduces the number of memory accesses to $n_p$, which is significant particularly for training data with low positive rate (in our case, $2 * n = 189364$ while $n_p = 2080$). Sorting the training instances by class is a cheap operation which is performed only once during initialization and can be done by the host CPU.

Counting of positive instances above the candidate's threshold is performed independently, in parallel, for each candidate. However, because there may not be enough candidates to fully occupy the GPU, we also split the counting process for each candidate into 256 independent sub-counts: the first sub-counter processes training instances 0, 256, 512, \ldots, the second sub-counter instances 1, 257, 513, \ldots, such that the sub-counts can be executed in parallel. The striped assignment of training instance indices to sub-counters is necessary to take advantage of the full memory bandwidth of the GPU. The host CPU then aggregates the sub-counts by summing, which is a very lightweight operation.

@@

% ############################################################################
\section{Conclusion} \label{concl}
% ############################################################################
We have presented an implementation of GA training of feedforward ANN classifiers for the CUDA platform for GPU programming. By carefully designing memory organization, algorithm computational load and memory access patterns, we have obtained a 10-fold speedup compared to a conventional sequential CPU implementation. A multipass approach was required for part of our system to compensate for the limited amount of fast on-chip GPU memory. Our method scales across population and training set sizes and is expected to be useful in other data intensive machine learning and data mining tasks.

>>>>

% ----------------------------------------------------------------------------
\subsection{Computation of Number of Truly Positive Top 20\% Instances} \label{truepos}
% ----------------------------------------------------------------------------
We wish to compute the number of truly positive instances in the top 20\% of scores. Assuming the training instances are unsorted, we must read from memory each instance's true class and corresponding ANN output value and compare the output value to the candidate's threshold. For $n$ training instances, this requires $2 * n$ memory accesses. However, by pre-sorting the training data such that the $n_p$ positive instances appear first in memory, we need only read the first $n_p$ ANN outputs, knowing that they correspond to the positive instances. This reduces the number of memory accesses to $n_p$, which is significant particularly for training data with low positive rate (in our case, $2 * n = 189364$ while $n_p = 2080$). Sorting the training instances by class is a cheap operation which is performed only once during initialization and can be done by the host CPU.

Counting of positive instances above the candidate's threshold is performed independently, in parallel, for each candidate ANN. However, because there may not be enough candidates to fully occupy the GPU, we also split the counting process for each candidate into 256 independent sub-counts: the first sub-counter processes training instances 0, 256, 512, \ldots, the second sub-counter instances 1, 257, 513, \ldots, such that the sub-counts can be executed in parallel. The striped assignment of training instance indices to sub-counters is necessary to take advantage of the full memory bandwidth of the GPU. The host CPU then aggregates the sub-counts by summing, which is a very lightweight operation.

@@

% ############################################################################
\section{Conclusion} \label{concl}
% ############################################################################
We have presented an implementation of GA training of feedforward ANN classifiers for the CUDA platform for GPU programming. By carefully designing memory organization, algorithm computational load and memory access patterns, we have obtained a 10-fold speedup compared to a conventional sequential CPU implementation. A multipass approach was required in the selection step of classifier evaluation to compensate for the limited amount of fast on-chip GPU memory. Our method scales across population and training set sizes and is expected to be useful in other data intensive machine learning and data mining tasks.
